peter_models

### 2024-12-07 peter_models.qmd
update models

# Fitting OLR Model

```{r}
dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
dataset_clean$Q271 <- factor(dataset_clean$Q271,
                            levels = c("(5) Not at all important",
                                     "(4) Slightly important",
                                     "(3) Moderately important", 
                                     "(2) Very important",
                                     "(1) Extremely important"),
                            ordered = TRUE)

@@

-   The nominal test for Model 1 shows significant violations (p \< 0.05 for all predictor variables except for Religion

@@

    -   Given the large sample size (17,545), this number of influential points is not extremely concerning.
2.  Key Findings from Model 1:
    1.  The different levels of racial identity importance are all significant
    2.  There are differences among racial/ethnic groups
3.  Predicted Relationships (from plot)
    1.  There appears to be a positive relationship between racial identity importance and political interest

# Fit multinomial model
m_mult <- multinom(Q29_unordered ~ Q271 + race_recoded + age_recoded +
            rural + # Rural or Urban/Suburban
            sex_orientation + # Sexual Orientation
            Religion + 
            educ + # Educational Attainment
            S3B + # Gender
            S7 + # Nativity
            Q21, # Party affiliation
          data = dataset_clean)

print("McFadden's R-squared:")
print(Basic = mcfadden_r2)

# Print AIC
print(Multinomial = AIC(m_mult))

# Fitting Partial Proportional Odds Model

```{r}
# Load required packages
library(VGAM)
library(ordinal)
library(tidyverse)
library(sandwich)

# Create ordered 3-category version

# Fit PPO model

# create prediction data

# get predictions for PPO model

# Reshape for plotting

# create plot

# Model summaries

# Cross-validation

# calculate variable importance for PPO model

# PPO Model with Interactions

>>>>

# Fitting OLR Model Using polr()

```{r}
library(MASS) # package for model
library(caret) # package for confusion matrix

dataset_clean$Q271_reordered <- factor(dataset_clean$Q271,
                            levels = c("(5) Not at all important",
                                     "(4) Slightly important",
                                     "(3) Moderately important",
                                     "(2) Very important",
                                     "(1) Extremely important"))

ord_mod <- polr(Q29 ~ Q271_reordered + race_recoded + age_recoded +
            rural + # Rural or Urban/Suburban
            sex_orientation + # Sexual Orientation
            educ + # Educational Attainment
            S3B + # Gender
            S7 + # Nativity
            Q21, # Party affiliation
          data = dataset_clean, Hess=TRUE)

summary(ord_mod)

pvals <- pnorm(-abs(summary(ord_mod)$coef[,"t value"]))*2
ctable <- cbind(summary(ord_mod)$coef,pvals)

ctable

(exp_coefs <- exp(cbind(OR=coef(ord_mod),confint(ord_mod))))

head(ord_mod$fitted.values)
head(predict(ord_mod))

confusionMatrix(predict(ord_mod), dataset_clean$Q29, mode="everything")

```{r}
# generate predictions for the ordinal model

new_dataset <- data.frame

predict(ord_mod, new_dataset)
```

# Fit multinomial model
m_mult <- multinom(Q29_unordered ~ Q271 + race_recoded + age_recoded +
            rural + # Rural or Urban/Suburban
            sex_orientation + # Sexual Orientation
            educ + # Educational Attainment
            S3B + # Gender
            S7 + # Nativity
            Q21, # Party affiliation
          data = dataset_clean)

print("McFadden's R-squared:")
print(mcfadden_r2)

# Print AIC
print(AIC(m_mult))

# Fitting OLR Model Using clm() (Following Jennifer's method)

```{r}
dataset_clean$Q271_simple <- as.numeric(dataset_clean$Q271)
# dataset_clean$Q271 <- factor(dataset_clean$Q271,
#                             levels = c("(5) Not at all important",
#                                      "(4) Slightly important",
#                                      "(3) Moderately important",
#                                      "(2) Very important",
#                                      "(1) Extremely important"),
#                             ordered = TRUE)

@@

-   The nominal test for Model 1 shows significant violations (p \< 0.05 for all predictor variables except for Religion)

@@

    -   Given the large sample size (17,545), this number of influential points is not extremely concerning.
2.  Key Findings from Model 1:
    1.  The different levels of racial identity importance are all significant
    2.  There are differences among racial/ethnic groups
3.  Predicted Relationships (from plot)
    1.
